---
title: "LMreg2"
author: "gc5k"
date: "3/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pROC) #Roc
library(MASS) #stepAIC
library(psych) #pairs, pairs.panel
library(glmnet) #lasso
library(lmvar) #cv.lm, does not allow missing data
library(boot) #cv.glm
library(randomForest) #
library(mice) #md.pattern
library(leaps) #best subset
library(varhandle) #numeric check:
source("~/R/MyLib/shotgun.R")

```

# COVID-19 {.tabset .tabset-fade .tabset-pills}

## Read data & QC
```{r readdata}
DatOrg=read.csv("./data2/covID2019zj.cgb1.csv", header = T)
#dat=dat[dat[,22]!=0.5,] #delete missing value
dat=DatOrg
md.na=md.pattern(dat) #check missing value pattern
rmNA=which(apply(is.na(dat), 1, any)) #remove
dat=DatOrg

if(length(rmNA) > 0) {
  dt=data.frame(dat[-rmNA,])
  print(paste("Remove", length(rmNA), "records due to missing."))
  print(paste(nrow(dt), "samples remained."))
} else {
  dt=dat
}

#check whether numeric
for(i in 2:ncol(dt)) {
  isNum=which(!varhandle::check.numeric(dt[,i]))
  if(length(isNum) > 0) {
    print(colnames(dt)[i])
    print(isNum)
  }
}

cl=which(apply(dt, 2, var)==0)
print(paste("removed", names(dt)[cl], "for no variation"))
dt=dt[,-1*unique(c(1, 37, cl))] #remove ID and DaiScore
```

## Feature selection
We evaluate ROC for each associated clinical measure. As seem below, it is nearly equivalent to logistic model analysis. We also conducted simple linear regression for "clinical score" that summarises all clinical measures.

```{r ROC, prompt=FALSE}
#feature selection using AUC
aucMat=matrix(0, ncol(dt)-1, 1)
rownames(aucMat)=colnames(dt)[1:nrow(aucMat)]
for(i in 1:nrow(aucMat)) {
  rocObj=roc(dt$diag, dt[,i], quiet = T)
  aucMat[i,1]=rocObj$auc
}

par(las=2)
layout(matrix(c(1,1,3,2,2,3), byrow=T, 2, 3))
barplot(main="COVID ROC", aucMat[,1], beside = T, border = F, col="green")

#######simple linear regression
respose=dt[,c("diag")]
fm=c("binomial")

sReg=array(1, dim=c(ncol(dt)-1, 7))
rownames(sReg)=colnames(dt)[-ncol(dt)]
for(j in 1:nrow(sReg)) {
    lm1=glm(respose~dt[,j], family=fm, data=dt)
    sReg[j,1:4]=summary(lm1)$coefficient[2,]
    sReg[j,5]=lm1$null.deviance
    sReg[j,6]=lm1$deviance
}

par(mar=c(6,4,2,2))
logP=-log10(sReg[,4])
barplot(logP, angle = 45, beside = T, horiz = F, border = F, ylab="-log10(p)", col=c("green"), main="Feature selection")
abline(h=c(8,10), col=c("red", "blue"), lty=2)
legend("topleft", legend = c("KitDiag"), pch=15, col=c("green"), bty = 'n')

plot(aucMat[,1], logP)
```

## glm Cross-validation
The logistic regression for case-control study is $log(\frac{p}{1-p})={\mathbf{X}\beta}+e$, in which $\mathbf{X}$ the matrix for predictors and $\beta$ is the vector for their effects. The logistic regression can be solved using Fisher scoring algorithm. Given the estimated $\beta$
Probability is $$p=\frac{1}{1+e^{X\hat{\beta}}}$$.

```{r glmR}
sink("cv.txt")
pcut=c(1, 3, 5, 10, 15)
Ksize=5
KF=sample(KFold(Ksize, nrow(dt)), nrow(dt))
aucFull=array(0, dim=length(pcut))
aucV=matrix(0, nrow=length(pcut), ncol=Ksize)

for(i in 1:length(pcut)) {
  xName=c(rownames(sReg)[which(-log10(sReg[,4])>pcut[i])], "diag")
#  print(xName)
  dtFull=dt[, xName]
  glmFull=glm(diag~., data=dtFull, family = "binomial")
  pre=exp(glmFull$fitted.values)/(1+exp(glmFull$fitted.values))
  aucObjFul=roc(dtFull$diag, pre, quiet=T)
  aucFull[i]=aucObjFul$auc
  print(paste("============Mode ", i))
  print(glmFull$coefficients)
  print(paste("Full model AUC:", aucObjFul$auc))
  print(ci(aucObjFul))
  for(j in 1:Ksize) {
    dtTr=dt[KF!=j, xName]
    dtT=dt[KF==j, xName]

    glmTr=glm(diag~., data=dtTr, family = "binomial")
    xP=cbind(matrix(rep(1, nrow(dtT)), nrow=nrow(dtT), 1),
             as.matrix(dtT[,-ncol(dtT)]))

    betaTr=matrix(glmTr$coefficients, ncol=1)
    preT=exp(xP%*%betaTr)/(1+exp(xP%*%betaTr))
    aucObjT=roc(dtT$diag, preT[,1], quiet=T)
    aucV[i,j] = aucObjT$auc
    
    print(paste("=Cross-validation ", j, "/", Ksize))
#  print(glmFull$coefficients)
    print(paste("Test model AUC:", aucObjT$auc))
    print(ci(aucObjT))
  }
  print("")
}

aucBd=cbind(aucFull, aucV)
rownames(aucBd)=pcut
barplot(t(aucBd), beside = T, xlab="-log10(p-cut)", border = F, col=c("red", rep("grey",Ksize)), ylim=c(0,1))
abline(h=0.9)
```

## Chosen cross-validation
$p=\frac{e^{\mathbf{X}\beta}}{1+e^{\mathbf{X}\beta}}$
```{r dai-glmR}
sink("dai-cv.txt")
Ksize=5
KF=sample(KFold(Ksize, nrow(dt)), nrow(dt))
aucFull=array(0, dim=1)
aucV=matrix(0, nrow=1, ncol=Ksize)

xName=c("TravelWuhan", "SickWuhan", "SickOther", "DiseaseCluster", "Fatigue", "ShortBreath", "SoreMuscle", "WhiteCellCnt", "CT", "diag")
for(i in 1:1) {
  dtFull=dt[, xName]
  glmFull=glm(diag~., data=dtFull, family = "binomial")
  pre=exp(glmFull$fitted.values)/(1+exp(glmFull$fitted.values))
  aucObjFul=roc(dtFull$diag, pre, quiet=T)
  aucFull[i]=aucObjFul$auc
  cat("============Full Mode ", i, "\n")
  print(summary(glmFull)$coefficients)
  cat(paste("Full model AUC:", aucObjFul$auc, "\n"))
  print(ci(aucObjFul))
  cat('\n\n')
  for(j in 1:Ksize) {
    print(paste("=Cross-validation ", j, "/", Ksize))
    dtTr=dt[KF!=j, xName]
    dtT=dt[KF==j, xName]

    glmTr=glm(diag~., data=dtTr, family = "binomial")
    print(summary(glmTr)$coefficients)
    xP=cbind(matrix(rep(1, nrow(dtT)), nrow=nrow(dtT), 1),
             as.matrix(dtT[,-ncol(dtT)]))

    betaTr=matrix(glmTr$coefficients, ncol=1)
    preT=exp(xP%*%betaTr)/(1+exp(xP%*%betaTr))
    aucObjT=roc(dtT$diag, preT[,1], quiet=T)
    aucV[i,j] = aucObjT$auc
    
#    print(glmFull$coefficients)
    cat(paste("Test model AUC:", aucObjT$auc, "\n"))
    print(ci(aucObjT))
    cat("\n\n")
  }
}

cat(paste("mean:", mean(aucV), "sd:", sd(aucV)))
aucBd=cbind(aucFull, aucV)
rownames(aucBd)="Full"
barplot(t(aucBd), beside = T, xlab="-log10(p-cut)", border = F, col=c("red", rep("grey",Ksize)), ylim=c(0,1))
abline(h=0.9)
sink()
```

## Session Information
```{r Rsys}
sessionInfo()
```

