---
title: "EigenGC"
author: "Chen Guo-Bo [chenguobo@gmail.com]"
date: "`r Sys.Date()`"
output:
 html_document:
    theme: united
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Table of contents {.tabset .tabset-fade .tabset-pills}
```{r}
print(Sys.time())
```
## Mixture distribution for many $\chi^2$
It is related to [Devlin&Roeder](https://github.com/gc5k/Notes/blob/master/Rpub/EigenGC/DevlinRoeder1999.pdf). In its Eq 7, it is 
$$Y^2 \sim  \epsilon \chi^2_{1,A^2}+(1-\epsilon) \chi^2_{1,0}$$
Here, it is generalized as

$$Y^2 \sim  \sum_{k=1}^K\chi^2_{1,ncp_k}$$

## NCP and $\lambda_{GC}$
$median(\chi^2_{1,ncp})=ncp$ when $ncp$ is smaller than 0.2, and $\lambda_{GC} \approx 1+ncp$

$median(\chi^2_{1,ncp})=ncp$ when $ncp$ is greater than 1, and $\lambda_{GC}=\frac{ncp}{\chi^2_{1, ncp=0}}=\frac{ncp}{0.455}$

```{r, gc-smallNCP}
layout(matrix(1:2, 1, 2))
NCPA=seq(0.01, 0.2, 0.01)
plot(main="Small NCP (<0.2)", NCPA, qchisq(0.5, 1, NCPA)/0.455, xlab="NCP", ylab="LambdaGC")
abline(a=1, b=1, col=c("blue"))
abline(a=1, b=2, col=c("red"), lty=2)
legend("topleft", legend = c("b=1", "b=2"), col=c("blue", "red"), lty=c(1,2))

NCPB=c(seq(0.01, 0.2, 0.01),seq(1, 10, 0.5))
plot(main="Large NCP (>1)",NCPB, qchisq(0.5, 1, NCPB)/0.455, xlab="NCP", ylab="LambdaGC")
abline(a=0, b=1, col=c("blue"))
abline(a=0, b=2, col=c("red"), lty=2)
legend("topleft", legend = c("b=1", "b=2"), col=c("blue", "red"), lty=c(1,2))

```

## Estimate NCP from $\lambda_{GC}$

```{r, NCP_GC}
NCP=c(0.1, 0.25, 0.5, 1.1, 1.25, 1.5)
SM=10000
for(i in 1:length(NCP)) {
  layout(matrix(1:2, 1, 2, byrow = T))
  dat=rchisq(SM, 1, ncp=NCP[i])
  gc=median(dat)/0.455

  qqplot(rchisq(SM, 1), dat/gc,main=paste0("NCP=", NCP[i]), pch=16, cex=0.5)
  abline(a=0, b=1, col="red")
  legend("topleft", legend = format(gc, digits = 4))
  qqplot(rchisq(SM,1, ncp=gc/2), dat, main=paste0("NCP=", NCP[i]), pch=16, cex=0.5)
  abline(a=0, b=1, col="blue")
  legend("topleft", legend = format(gc/2, digits=4))
}
```

## Simulated Example 1 (frq=0.5)
```{r, two-group}
# lambda1 mixture distribution / genetic drift only
M=10000
f=0.01 # genetic drift

P=runif(M, 0.5, 0.5) # ancestral
Sz=c(300, 300) #subgroup sample size
Fq=matrix(0, 2, M) 
Fq[1,]=rbeta(M, P*(1-f)/f, (1-P)*(1-f)/f)
Fq[2,]=rbeta(M, P*(1-f)/f, (1-P)*(1-f)/f)

G=matrix(0, sum(Sz), M)
for(i in 1:Sz[1]) {
  G[i,] = rbinom(M, 2, Fq[1,])
}

for(i in (Sz[1]+1):(Sz[1]+Sz[2])) {
  G[i,] = rbinom(M, 2, Fq[2,])
}

fq1 = colMeans(G[1:Sz[1],])/2 # observed allele freq, in subgroup 1
fq2 = colMeans(G[(Sz[1]+1):(Sz[1]+Sz[2]),])/2 # observed allele freq, in subgroup 2
H1 = 2*fq1*(1-fq1)
H2 = 2*fq2*(1-fq2)
Hs = (H1+H2)/2
HT = 2*((fq1+fq2)/2)*(1-(fq1+fq2)/2)
FstL = (fq1-fq2)^2/(HT)*(Sz[1]+Sz[2])
Fst = mean((fq1-fq2)^2/(HT))*(Sz[1]+Sz[2])

Gs=apply(G, 2, scale)
GG=Gs %*% t(Gs) / M
EigenG=eigen(GG)
layout(matrix(1:2, 1, 2))
barplot(EigenG$values[1:20])
plot(main=paste(M, "markers"), EigenG$vectors[,1], EigenG$vectors[,2], bty='n', xlab="PC 1", ylab="PC 2", pch=16, col=c(rep("red", Sz[1]), rep("blue", Sz[2])))

# get lambda GC
eigenspace1 = EigenG$vectors[,1]
para = matrix(0, M, 2)
for (i in 1:M){
  model = lm(eigenspace1~Gs[,i])
  para[i,1]=summary(model)$coefficients[2,3]^2 #p value
  para[i,2]=summary(model)$coefficients[2,4] #p value
}
lambdaGC = qchisq(median(para[,2]), 1, lower.tail = F)/qchisq(0.5, 1, lower.tail = F)
lambda1 = EigenG$values[1]
layout(matrix(1:1, 1))
a = matrix(c(Fst, lambda1, lambdaGC), 1, 3)
colnames(a)=c("Fst", "Eigenvalue 1", "LambdaGC")
barplot(a)

layout(matrix(1:2, 1, 2))
qqplot(rchisq(nrow(para), 1, ncp=Fst), para[,1])
qqplot(rchisq(nrow(para), 1, ncp=FstL), para[,1])

```

## Simulated Example 2 (frq=0.2~0.8)
```{r, two-group-het}
# lambda1 mixture distribution / genetic drift only
M=10000
f=0.01 # genetic drift

P=runif(M, 0.2, 0.8) # ancestral
Sz=c(300, 300) #subgroup sample size
Fq=matrix(0, 2, M) 
Fq[1,]=rbeta(M, P*(1-f)/f, (1-P)*(1-f)/f)
Fq[2,]=rbeta(M, P*(1-f)/f, (1-P)*(1-f)/f)

G=matrix(0, sum(Sz), M)
for(i in 1:Sz[1]) {
  G[i,] = rbinom(M, 2, Fq[1,])
}

for(i in (Sz[1]+1):(Sz[1]+Sz[2])) {
  G[i,] = rbinom(M, 2, Fq[2,])
}

fq1 = colMeans(G[1:Sz[1],])/2 # observed allele freq, in subgroup 1
fq2 = colMeans(G[(Sz[1]+1):(Sz[1]+Sz[2]),])/2 # observed allele freq, in subgroup 2
H1 = 2*fq1*(1-fq1)
H2 = 2*fq2*(1-fq2)
Hs = (H1+H2)/2
HT = 2*((fq1+fq2)/2)*(1-(fq1+fq2)/2)
FstL = (fq1-fq2)^2/(HT)*(Sz[1]+Sz[2])
Fst = mean((fq1-fq2)^2/(HT))*(Sz[1]+Sz[2])

Gs=apply(G, 2, scale)
GG=Gs %*% t(Gs) / M
EigenG=eigen(GG)
layout(matrix(1:2, 1, 2))
barplot(EigenG$values[1:20])
plot(main=paste(M, "markers"), EigenG$vectors[,1], EigenG$vectors[,2], bty='n', xlab="PC 1", ylab="PC 2", pch=16, col=c(rep("red", Sz[1]), rep("blue", Sz[2])))

# get lambda GC
eigenspace1 = EigenG$vectors[,1]
para = matrix(0, M, 2)
for (i in 1:M){
  model = lm(eigenspace1~Gs[,i])
  para[i,1]=summary(model)$coefficients[2,3]^2 #p value
  para[i,2]=summary(model)$coefficients[2,4] #p value
}
lambdaGC = qchisq(median(para[,2]), 1, lower.tail = F)/qchisq(0.5, 1, lower.tail = F)
lambda1 = EigenG$values[1]
layout(matrix(1:1, 1))
a = matrix(c(Fst, lambda1, lambdaGC), 1, 3)
colnames(a)=c("Fst", "Eigenvalue 1", "LambdaGC")
barplot(a)

layout(matrix(1:2, 1, 2))
qqplot(rchisq(nrow(para), 1, ncp=Fst), para[,1])
qqplot(rchisq(nrow(para), 1, ncp=FstL), para[,1])

```

## Arabdopsis Eigen1
Arabidopsis has 295 lines, and 150K markers. Eigenvalue 1=9.4, $\lambda_{GC}=9.3$
```{r, arabdopsis-1}
conn <- gzfile("arab.1.egwas.gz", "rt")
eg <- read.table(conn, as.is = T, header = T)
eg$P0=pchisq(eg$Chi, 1, lower.tail = F)
library(qqman)

print(paste("GC correction"))
gc=median(eg$Chi)/qchisq(0.5, 1)
eg$P0gc=pchisq(eg$Chi/gc, 1, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0gc")

print(paste("ncp=gc correction"))
eg$P0ncp=pchisq(eg$Chi, 1, ncp=gc, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncp")

print(paste("ncp=mean(fst)*N"))
eg$P0ncpFst=pchisq(eg$Chi, 1, ncp=mean(eg$Fst)*295, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncpFst")

hist(eg$Fst)
print(paste("ncp=mean(fst)+sd(fst)*1.96"))
fstMk=mean(eg$Fst)+sd(eg$Fst)*1.96
eg$P0ncpFstH=pchisq(eg$Chi, 1, ncp=fstMk*295, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncpFstH")
```


## Arabdopsis Eigen2

```{r, arabdopsis-2}
conn <- gzfile("arab.2.egwas.gz", "rt")
eg <- read.table(conn, as.is = T, header = T)
eg$P0=pchisq(eg$Chi, 1, lower.tail = F)
library(qqman)

print(paste("GC correction"))
gc=median(eg$Chi)/qchisq(0.5, 1)
eg$P0gc=pchisq(eg$Chi/gc, 1, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0gc")

print(paste("ncp=gc correction"))
eg$P0ncp=pchisq(eg$Chi, 1, ncp=gc, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncp")

print(paste("ncp=mean(fst)*N"))
eg$P0ncpFst=pchisq(eg$Chi, 1, ncp=mean(eg$Fst)*295, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncpFst")

hist(eg$Fst)
print(paste("ncp=mean(fst)+sd(fst)*1.96"))
fstMk=mean(eg$Fst)+sd(eg$Fst)*1.96
eg$P0ncpFstH=pchisq(eg$Chi, 1, ncp=fstMk*295, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncpFstH")

```


## CEU vs TSI
$\lambda_{GC}=1.708$
```{r, euro}
conn <- gzfile("eu.1.egwas.gz", "rt")
eg <- read.table(conn, as.is = T, header = T)
eg$P0=pchisq(eg$Chi, 1, lower.tail = F)
library(qqman)

print(paste("GC correction"))
gc=median(eg$Chi)/qchisq(0.5, 1)
eg$P0gc=pchisq(eg$Chi/gc, 1, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0gc")

print(paste("ncp=gc correction"))
eg$P0ncp=pchisq(eg$Chi, 1, ncp=gc, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncp")

print(paste("ncp=mean(fst)*N"))
eg$P0ncpFst=pchisq(eg$Chi, 1, ncp=mean(eg$Fst)*295, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncpFst")

hist(eg$Fst)
print(paste("ncp=mean(fst)+sd(fst)*1.96"))
fstMk=mean(eg$Fst)+sd(eg$Fst)*1.96
eg$P0ncpFstH=pchisq(eg$Chi, 1, ncp=fstMk*295, lower.tail = F)
manhattan(eg, chr="CHR", bp="BP", p="P0ncpFstH")

```

## R.sys

```{r, R.sys}
sessionInfo()
```
